\documentclass[usegeometry,parskip=half]{scrartcl}

\usepackage[
  a4paper,
  top=25mm,
  bottom=25mm,
  left=25mm,
  right=25mm,
  footskip=10mm
]{geometry}

\usepackage[hidelinks]{hyperref}
\pagenumbering{gobble}

\usepackage{microtype}
\usepackage{fontspec}
\usepackage{unicode-math}

\setmainfont{Stix Two Text}
\setsansfont{TeX Gyre Heros}
\setmathfont{Stix Two Math} 

\begin{document}
\textbf{ICS0036 Cryptography --- evaluation criteria}

Your final grade is determined as follows:
\begin{table}[h]
  \centering
  \begin{tabular}{ccc}
    40\% - homework assignments & 10\% - group presentation & 50\% - final exam
  \end{tabular}
\end{table}

To pass (i.e. get a non-zero grade), you must satisfy the following:
\begin{enumerate}
  \item You must cumulatively score at least 50\% (20/40) on your homework assignments.
  \begin{itemize}
    \item There will be at most $6$ homework assignments.
    \item You will have at least $2$ weeks to submit each homework.
    \item If you cannot meet the deadline, write to me no later than 4 days before the deadline with your reason.
    I will not grant you an extension if you miss the notice period.
    \item When you are stuck:
    \begin{itemize}
      \item You may turn to me for help after classes/by email/by Discord.
      \item You may give pointers to each other, but not share solutions.
    \end{itemize}
    \item Grace clause: if I find mistakes in your homework, you will have one week to fix them to gain back up to half of the lost points.
    E.g. for a homework originally graded 7/10, you can fix it to get up to 8.5/10.
    \item Missed deadline: you can submit a homework up to one week late for half of the grade, with no grace clause.
    Submitting more than a week late: 0.
  \end{itemize}
  \item {[Attend 8 lectures AND 8 practices] OR [face the stricter exam pass criteria].}
  \item {[Attend the midterm] AND [pass your in-class presentation ($\ge$ 5/10)].}
  \begin{itemize}
    \item The midterm is a mandatory but ungraded, consider it a `reality check' for your knowledge.
    \item 10min group presentation (2 or 3 people).
    \item Each group member must be able to answer my questions.
    Individual failure is therefore possible, so make sure that everyone contributes.
  \end{itemize}
  \item Attend and pass the in-person written exam.
  \begin{itemize}
    \item 1h 30min closed book exam with a handwritten one-sided A4 cheat-sheet allowed.
    \item Alternatively: 20min oral exam (max 10 students).
    \item You must pass \emph{each} of the sections in the exam.
    The pass threshold for each section will be 2$\times$ higher for those with insufficient attendance, e.g. instead of needing 2p per section, you will need 4p per section.
    \item 1 retake allowed.
    \item All previous requirements must be satisfied to attend.
  \end{itemize}
\end{enumerate}

The reasons for the above criteria and for the closed book exam are explained in the intro lecture.

\vspace*{\parskip}

Violation of good academic practice:

Copying someone's homework, plagiarism, etc.\footnotemark{}, fails you the course and I will report you.

\footnotetext{Chapter 9 of TalTech's academic policies: \url{https://taltech.ee/en/academic-policies-taltech}}
\end{document}
